---
title: 'Class5: Data Visualization with GGPLOT'
author: "Ashley Vang"
date: "10/11/2022"
output:
  pdf_document: default
  html_document: default
---

#Our first ggplot

To use the ggplot2 package I first need to have it installed on my computer. 

To install any package we use the 'install.packages()' command 

Now can I use it? NO! first we need to call 'library(ggplot2)' 

```{r}
library(ggplot2)
ggplot()
```

```{r}
mpg
```

Our first plot of displ vs hwy 
All ggplot() graphs are made in the same way 

data + aes + geoms 

```{r}
ggplot(mpg) + 
  aes(x=displ, y=hwy) + 
  geom_point()
```

I can add more layers 

```{r}
ggplot(mpg) + 
  aes(x=displ, y=hwy) + 
  geom_point() +
  geom_smooth(se = FALSE, method = lm)
```

# Plot of gene expression data 

First read the data from online. 

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

> Q. How many genes are in this dataset? 

```{r}
nrow(genes)
```

What are the colnames?
```{r}
colnames(genes)
```

A first version plot of this data Condition1 vs Condition2
```{r}
ggplot(genes) +
  aes(x=Condition1, y=Condition2) + 
  geom_point()
```

Let's add some color. To do this we can add another aes() mapping of color to the 'State' column in our data. 

```{r}
ggplot(genes) +
  aes(x=Condition1, y=Condition2, col=State) + 
  geom_point() 
```

> Q. How many genes are up regulated and down regulated. 

```{r}
head(genes)
```

To get at just the State column 
```{r}
table (genes$State)
```

```{r}
p <- ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point()
```

We can add more layers 
```{r}
p +scale_color_manual(values = (c ("blue", "gray", "red"))) +
  labs(title="Gene Expression",
   x= "No drug", y="Drug",
   caption= "Made by me")
```

Applying to geom_point applies it to the entire plot, aes is specific to the data set 
```{r}
 ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point(alpha=0.3)
```
p

