---
title: "Homework 6"
author: "Ashley Vang"
date: "10/16/2022"
output: html_document
---

#Original Code
# Can you improve this analysis code?
library(bio3d)

s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")



We can see some errors in the code which we can fix. 
Change the s1 from the s3.b code into s3. 
```{r}
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

Notice how the code has elements that are specific and can be turned into a singular function such as read.pdb, trim.pdb, s.chainA$atom$b, plotb3 

Set repeats as specific names so we can input it easier into the functions. 
Set x <- to any protein in the pdb 
```{r}
x<- "specific protein"
x.chainA <- trim.pdb(s1, chain="A", elety="CA")

c<- x.chainA 
x.b <- c$atom$b 


plotb3(x.b,sse=x.chainA, typ="l", ylab="Bfactor")

```


Then input into a function using the three elements of a written function. 

Refresher:All functions in R consist of at least 3 things: 

-A **name** (we can pick this but it must start with a character)
-Input **arguments** (there can be multiple comma separated inputs)
-The **body **(where work actually happens)

```{r}
protein <- function(x) { x.chainA <- trim.pdb (x, chain="A", elety="CA")
c<- x.chainA 
x.b <- c$atom$b 

plotb3(x.b,sse=x.chainA, typ="l", ylab="Bfactor")}
```

Test to see if it works...
```{r}
protein(s1)
```

```{r}
protein(s2)
```

```{r}
protein(s3)
```

Function protein encodes a code that takes any protein from the pdb and creates a plot of the ratio of Bfactor to residue. 
To use the function just write protein() and insert your protein of interest. 



