---
title: 'Class 10: Halloween Mini Project'
author: "Ashley Vang"
date: "10/27/2022"
output:
  html_document: default
  pdf_document: default
---
## Importing Candy Data

First save the input into your project directory. 

```{r}
candy_file <- "candy-data.csv"
```

Then load and input the data.
```{r}
candy= read.csv(candy_file, row.names=1)
head(candy)

```

>Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```
There are 85 different types of candy. 


>Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)

```

There are 38 fruity candy types in the dataset. 

We can find the winpercent value for Twix. (Those with a higher value people like more)

## What is your favorite candy?

```{r}
rownames(candy)
```


```{r}
candy["Twix", ]$winpercent
```


>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Sour Patch Kids",]$winpercent
```


My favorite candy is sour patch kids. Its winpercent value is 59.864. 


>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",]$winpercent
```
The winpercent for kit kat is 76.7686. 


>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```
The winpercent value for tootsie roll snack bars is 49.6535 


We can look at the overview of the given dataset. 

```{r}
library("skimr")
skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

The histogram column that plots the winpercent looks to be on a different scale. It is not on a 0 to 1 scale or out of 100 scale, it is not on a numerical scale. 

>Q7. What do you think a zero and one represent for the candy$chocolate column?

```{r}
candy$chocolate
```

A 1 means there is chocolate in that candy and a 0 means there is no chocolate in the candy. 

>Q8. Plot a histogram of winpercent values

We can plot using two different methods. hist() or ggplot().

```{r}
hist(candy$winpercent)
```



```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent) +
  geom_histogram(bins=10, col="red", fill="orange")
  
```


>Q9. Is the distribution of winpercent values symmetrical?

No, the distribution is not symmetrical, you can see the values are unevenly distributed. 

>Q10. Is the center of the distribution above or below 50%?

The center of distribution is above 50%.


>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}

chocolate.inds <-as.logical(candy$chocolate)
chocolatewins <-candy[chocolate.inds,]$winpercent
mean(candy[chocolate.inds,]$winpercent)


fruity.inds <- as.logical(candy$fruity)
fruitwins <-candy[fruity.inds,]$winpercent
mean(candy[fruity.inds,]$winpercent)
```
On average the chocolate candy is higher ranked than the fruity candy. 

>Q12. Is this difference statistically significant?

```{r}
t.test(chocolatewins, fruitwins)
```

There is a statistically significant difference between the two as seen from the low p-value. 


## Overall Candy Rankings


>Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```

The five least liked candy types are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters. 


>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),], n=5)
```
The top 5 favorite candy types are Snickers, Kit kats, twix, ReeseOs minis, and ReeseOs Peanut Butter Cup. 


>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent,rownames(candy)) +
  geom_col()
```

>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```




```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "yellow"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
#my_cols
```


```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```


>Q17. What is the worst ranked chocolate candy?

The worst ranked chocolate candy is sixlets. 

> Q18. What is the best ranked fruity candy?

The best ranked fruity candy is starburst. 



## Taking a look at pricepercent

We can plot the value for money per each candy. To see which is best in terms of price and win percent. 

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 10)
```
>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

The highest rank in terms of winpercent for the least money is Reese O's minis. 


>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

The top 5 most expensive are Nik L Nip, Nestle Smarties, Ring pop, HersheyOs Krackel, and HersheyOs Milk Chocolate. The least popular amongst these are Nik L Nip. 


>Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().

```{r}
# Make a lollipop chart of pricepercent
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```


## Exploring the correlation structure

We will plot the correlation by using corrplot().

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

The two variables that are anti-correlated are shown by the red colored dot which are chocolate and fruity. 

>Q23. Similarly, what two variables are most positively correlated?

Variables most positively correlated are winpercent (popularity) and chocolate. 


## Principal Component Analysis 

Lets apply PCA. 

```{r}
pca <-  prcomp(candy, scale=TRUE)
summary(pca)
```

Now we can plot our PCA score plot, PC1 vs PC2. 

```{r}
plot(pca$x[,1:2])
```


```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

We can also plot using ggplot. 

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

We can use ggrepel. 

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```
We can generate an interactive plot. 

```{r}
library(plotly)
```

```{r}
ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

The original variables picked up by PC1 in the positive direction are fruity, hard, and pluribus. Yes, it makes sense because thats what most fruity candies are and fall under. 